---
title: "Tables amd Visualizations"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(rvest)
library(plotly)
library(flextable)
library(stringr)
library(MetBrewer)
library(data.table)
library(crosstalk)
library(conflicted)

# Set session-wide pref for library conflicts
conflicts_prefer(dplyr::filter())
```

```{r echo=FALSE, message=FALSE, include = FALSE, warning=FALSE}

# Function for removing extra characters
remove_bullish <- function(bullish_cols) {
  as.numeric(gsub(
    pattern = "[^0-9.-]",
    replacement = "",
    x = bullish_cols
  ))
}

# Function for removing spaces and space-like characters 
remove_spaces <- function(spacey_cols) {
  gsub(
    pattern = "\\s+",
    replacement = " ",
    x = spacey_cols
  )
}



# Gather and clean data ----
# Define root directory for TBS website with salary data
url <-
  "https://www.canada.ca/en/treasury-board-secretariat/services/innovation/human-resources-statistics/diversity-inclusion-statistics/distribution-public-service-canada-employees-designated-sub-group-salary-range-"

# List of years to get data for 
years <- c("2023", "2022", "2021", "2020", "2019", "2018", "2017")
    

# Create function to get salary data from website over 5 years

get_salary_data <- function(Year) {
  
  # Progress message 
  message(paste("Webscraping", Year, "data."))
  
  # Match years to their index in the data
  year_dictionary <- data.frame(year = years,
                                index = c(1:length(years)))
  
  # Index for year
  year_index <-
    as.numeric(year_dictionary$index[year_dictionary$year == Year])
  
  # Pull dataset with rvest
  
  assign(paste("salary_df", # Create df's for each year
               Year, 
               sep = "_"), {
    merge(x = {
    
      ## VM and Black employees ----
    {html_nodes(read_html(paste0( 
      url,
      "members-visible-minorities")), 
      "table") |> 
      html_table(fill = TRUE)
    }[[year_index]][c(1:3, 25)] |> # remove unneeded cols
     # retrieve all employees and total vm column
    dplyr::rename( # fix col names
      salary_range = "Salary range ($)",
      count_all = "All employees",
      count_black = "Black",
      count_vm = "Members of visible minorities")},
  
        y = {
  
    # Indigenous employees----
    {html_nodes(read_html(paste0(url, 
                                "indigenous-peoples")), 
               "table") |>
      html_table(fill = TRUE)
    }[[year_index]][c(1, 11)] |> # retrieve total indigenous column
      dplyr::rename(salary_range = "Salary range ($)",
                    count_indigenous = "Indigenous Peoples")   },
    by = "salary_range",
    all = TRUE) |> 
    
    merge(y = {
  
    ## Persons with disabilities ----
    {html_nodes(read_html(paste0(url,
                                "persons-disabilities")), "table") |>
      html_table(fill = TRUE)
    }[[year_index]][c(1, 15)] |> # retrieve total pwd column
      dplyr::rename(salary_range = "Salary range ($)",
                    count_pwd = "Persons with disabilities")
    
    },
    by = "salary_range",
    all = TRUE) |>

    # Remove rows with salary or table 
  filter(!grepl("^Salary|Table", salary_range)) |> 
    mutate_at(vars(matches("count")), remove_bullish) |> # remove extra characters from counts
    mutate_at(vars(matches("salary")), remove_spaces) |> # remove extra spaces from counts
    filter(!grepl("Notes", salary_range))
  },
  envir = globalenv())
  
}




# Get data for all years
# This step usually takes a while to run due to accessing the API

# Record start time
start.time <- Sys.time()

# Pull salary data for all years
# Run time: ~ 
lapply(years, get_salary_data)  |>
  bind_rows()

# Record end time
end.time <- Sys.time()

# Print elapse time
time.taken <- end.time - start.time
message(paste("Elaspsed time of", time.taken , "mins.", sep = " "))

  
# list of name of databases by year
years_data_cat <- paste("salary_df",
                        years,
                        sep = "_") |>
  
  # Concat every year's data into one df
  mget() |>
  bind_rows(.id = "Year") |>
  
  # Make a year column 
  mutate(Year = substr(Year,
                       11,
                       14))  |>
  mutate(salary_range = 
           ifelse(grepl('Total', 
                        salary_range), 
                  "Total", 
                  salary_range)) |>
  
  # Merge any rows with "Total" in name by year
  group_by(salary_range, Year) |>
  summarise_all(list(~ sum(., na.rm = TRUE)))

# Write intermediate dataset if needed
# |> 
   # write.csv(file = "years_data_cat.csv")

# Read in csv version
# years_data_cat <- read.csv("years_data_cat.csv")



# Imputation: add in imputed values to replace suppressed numbers 
imputation_data <- read.csv("imputation_estimates.csv", encoding = "UTF-8")

# Remove non-space characters and replace with spaces 
imputation_data$salary_range <- str_replace_all(imputation_data$salary_range, "\\s", " ")

# Loop to impute numbers into df
for (i in 1:nrow(imputation_data)){
  # separate imputation date into groups
  imputation_data_black <- imputation_data |> 
    filter(Group == "count_black")
  imputation_data_indigenous <- imputation_data |> 
    filter(Group == "count_indigenous")
  imputation_data_vm <- imputation_data |> 
    filter(Group == "count_vm")
  # Use case_when to meet conditions from imputation data
  years_data_cat <- years_data_cat |> 
    mutate(old_val = count_black) |>
    mutate(count_black = case_when(
        Year == imputation_data_black$Year[i] & 
          salary_range == imputation_data_black$salary_range[i] ~  imputation_data_black$value[i], 
        .default = old_val
      )
    ) |>
    mutate(old_val = count_indigenous) |>
    mutate(count_indigenous = case_when(
        Year == imputation_data_indigenous$Year[i] & 
          salary_range == imputation_data_indigenous$salary_range[i] ~  imputation_data_indigenous$value[i], 
        .default = old_val
      )
    ) |>
        mutate(old_val = count_vm) |>
    mutate(count_vm = case_when(
        Year == imputation_data_vm$Year[i] & 
          salary_range == imputation_data_vm$salary_range[i] ~  imputation_data_vm$value[i], 
        .default = old_val
      )
    ) |>
    select(-c(old_val)) |>
             mutate(count_non_vm = count_all - count_vm) # calculate non vm (all - vm's)
}

# Fix 2022 data - collapse under 50K 
# 2022 data has both "Under 25K" and "Under 50K" salaries. Here we will collapse those two categories as well as the  "25K-49K" category into "Under 50K" for consistency 
years_data_cat <-
  years_data_cat |> mutate(salary_range = case_when(
    salary_range %in% c("Under 25,000",
                        "Under 50,000",
                        "25,000 to 49,999") ~ "Under 50,000",
    TRUE ~ salary_range
  )) |>
  group_by(salary_range, Year) |>
  summarise(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE), 
    count_non_vm = count_all - count_vm
  )


# Clear workspace of unneeded df's and functions in saved memory
rm(salary_df_2017, salary_df_2018, salary_df_2019, salary_df_2020, salary_df_2021, salary_df_2022, url, get_salary_data, remove_bullish, remove_spaces)

# read in salary bin data
temp <-read.csv("ee_dv_salarybins.csv", fileEncoding = "UTF-8-BOM")

# df for renaming cols
median_colnames <- c("salary_range", 
                       "Year", 
                       "count_all", 
                       "count_black", 
                       "count_vm", 
                       "count_indigenous", 
                       "count_pwd", 
                       "count_non_vm",
                        "bin_width",
                       "bin_width_cumsum", 
                       "cumsum_all",
                       "cumsum_black",
                       "cumsum_vm",
                       "cumsum_indigenous",
                       "cumsum_pwd",
                       "cumsum_non_vm",
                       "half_pop_all",
                       "half_pop_black",
                       "half_pop_vm",
                       "half_pop_indigenous",
                       "half_pop_pwd",
                       "half_pop_non_vm")
  
# Create final datasets with cumulative frequencies and bins
df <-
  merge(
    x = years_data_cat, # read in yearly salary data by group
    y = { # match dataset to bin categories and widths from different years
        bind_rows(
      {temp[, c(1:2)] |> dplyr::rename(bins = bins_2022, bin_width = bin_width_2022)},           
    {temp[, c(3:4)] |> dplyr::rename(bins = bins_2021, bin_width = bin_width_2021)}, 
    {temp[, c(5:6)] |> dplyr::rename(bins = bins_2020, bin_width = bin_width_2020)},
    {temp[, c(7:8)] |> dplyr::rename(bins = bins_2019, bin_width = bin_width_2019)},
    {temp[, c(9:10)] |> dplyr::rename(bins = bins_2018, bin_width = bin_width_2018)},
    {temp[, c(11:12)] |> dplyr::rename(bins = bins_2017, bin_width = bin_width_2017)},
    .id = "key"
  ) |>
  mutate(year = case_when( # add year "keys" (html pull indexes years by numbers)
    key == "1" ~ 2022,
    key == "2" ~ 2021,
    key == "3" ~ 2020,
    key == "4" ~ 2019,
    key == "5" ~ 2018, 
    key == "6" ~ 2017
  )) |> 
  select(-key) |>  # remove key data
  filter(!is.na(bins))
    },
    by.x = c("salary_range", "Year"),
    by.y = c("bins", "year"),
    all.x = TRUE
  )  |> arrange( # arrange salary ranges from lowest to highest using factors
    Year,
    factor(
      salary_range,
      levels = c(
        "Under 15,000",
        "Under 25,000",
        "Under 50,000",
        "15,000 to 49,999",
        "25,000 to 49,999",
        "50,000 to 54,999",
        "55,000 to 59,999",
        "60,000 to 64,999",
        "65,000 to 69,999",
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999",
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999",
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "100,000 and over",
        "250,000 and over",
        "Total"
      )
  )) |> 
  group_by(Year) |> # get cumulative frequencies with cumsum()
    mutate(
      cumsum_bin_width = cumsum(bin_width), 
      cumsum_all = cumsum(count_all),
      cumsum_black = cumsum(count_black),
      cumsum_vm = cumsum(count_vm),
      cumsum_indigenous = cumsum(count_indigenous),
      cumsum_pwd = cumsum(count_pwd),
      cumsum_non_vm = cumsum(count_non_vm)
    ) |> ungroup() |>
  pivot_longer( # change data to longform
    cols = c(starts_with("count"), starts_with("cumsum")),
    names_to = "group",
    values_to = "count_cumsum"
  ) |>
  group_by(Year, group) |> 
    mutate(total = ifelse((grepl(pattern = "count_", x = group) & salary_range == "Total"), count_cumsum, NA)) |> # if total, make cumsum NA
  fill(total, .direction = "up") |> 
  ungroup() |> 
  mutate(half_pop = total/2) |> # get half population per bin
  group_by(Year, group)

```

```{r echo=FALSE, message=FALSE, include = FALSE, warning=FALSE}

# Median Calculations -----------------------------------------------------

# Create a function that calculates median for each group for a year
  # median = lower limit of median group + ((half pop - cum freq up to median)/ pop of median group) * width of median group


est_median <- function(year, ee_group) {
  # pull salary data by year and change back to wide format
  median_df <-
    df |>
    filter(Year == year) |> # filter for selected year
    pivot_wider(names_from = "group", # change back to wide format
                values_from = c("count_cumsum", 
                                "total", 
                                "half_pop")) |>
     dplyr::select(-contains(c("total_cumsum", # remove unneeded cols
                               "total_count",
                               "half_pop_cumsum"))) 
  
  # rename cols
  median_colnames <- c("salary_range", 
                       "Year", 
                       "bin_width",
                       "count_all", 
                       "count_black", 
                       "count_vm", 
                       "count_indigenous", 
                       "count_pwd", 
                       "count_non_vm",
                       "bin_width_cumsum", 
                       "cumsum_all",
                       "cumsum_black",
                       "cumsum_vm",
                       "cumsum_indigenous",
                       "cumsum_pwd",
                       "cumsum_non_vm",
                       "half_pop_all",
                       "half_pop_black",
                       "half_pop_vm",
                       "half_pop_indigenous",
                       "half_pop_pwd",
                       "half_pop_non_vm")
  
  colnames(median_df) <- median_colnames
  half_pop <- as.numeric(first(median_df[paste("half_pop_", ee_group, sep = "")])[1])
    
  # find lower limit of median group
  median_group_ll <- max(median_df$bin_width_cumsum[median_df[paste("cumsum_", ee_group, sep = "")] < half_pop], na.rm=TRUE)
  
  # find the cumulative frequncy up the median
  cumsum_uptomedian <- max(median_df[paste("cumsum_", ee_group, sep = "")][median_df[paste("cumsum_", ee_group, sep = "")]< half_pop], na.rm=TRUE)
  
  # find the frequency of the median group
  freq_mediangroup <- median_df[paste("count_", ee_group, sep = "")][median_df[paste("cumsum_", ee_group, sep = "")] > half_pop][1]
  
  # find the width of the median group 
  width_mediangroup <- median_df$bin_width[median_df[paste("cumsum_", ee_group, sep = "")] > half_pop][1]
  
  # calculate median [finally (●'◡'●)]
  median <- median_group_ll + ((half_pop - cumsum_uptomedian ) / freq_mediangroup) * width_mediangroup  
  
  # Create output and place in global environment
  assign("median_table", 
         {data.table(med = median, 
                             med_year = year, 
                             group = ee_group)}, 
         envir=globalenv())
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# run across all years and groups
groups <- c("black", "vm", "indigenous", "pwd", "non_vm", "all")

# Create empty df's 
median_df_2022 <- data.frame(median = NA, med_year = NA, group = NA)
median_df_2021 <- data.frame(median = NA, med_year = NA, group = NA)
median_df_2020 <- data.frame(median = NA, med_year = NA, group = NA)
median_df_2019 <- data.frame(median = NA, med_year = NA, group = NA) 
median_df_2018 <- data.frame(median = NA, med_year = NA, group = NA)
median_df_2017 <- data.frame(median = NA, med_year = NA, group = NA)

# Get median for all years and groups
for (i in 1:length(groups)){
  median_df_2022[i,] <- est_median("2022", groups[i])
}
for (i in 1:length(groups)){
  median_df_2021[i,] <- est_median("2021", groups[i])
}
for (i in 1:length(groups)){
  median_df_2020[i,] <- est_median("2020", groups[i])
}
for (i in 1:length(groups)){
  median_df_2019[i,] <- est_median("2019", groups[i])
}
for (i in 1:length(groups)){
  median_df_2018[i,] <- est_median("2018", groups[i])
}
for (i in 1:length(groups)){
  median_df_2017[i,] <- est_median("2017", groups[i])
}
# Concat all median df's together with rbind
median_df_final <- rbind(median_df_2022, median_df_2021, median_df_2020, median_df_2019, median_df_2018, median_df_2017)
# Remove unneeded df's from memory
rm(median_df_2022, median_df_2021, median_df_2020, median_df_2019, median_df_2018, median_df_2017)

# Create function and apply to round to nearest $50
round_any = function(x, accuracy, f=round){f(x/ accuracy) * accuracy}
median_df_final$median <- round_any(median_df_final$median, accuracy = 50)

# Swap to wide format for presentation
median_df_wide <- spread(median_df_final, group, median) |> 
  select(med_year, all, non_vm, vm, black, indigenous, pwd) |> 
  arrange(desc(med_year))


```

::: {.content-visible when-profile="english"}
## Median Salary Table

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Add better group labels 
colnames(median_df_wide) <-
  c(
    "Year",
    "All Employees",
    "Non-Visible Minority Employees",
    "Visible Minority Employees",
    "Black Employees",
    "Indigenous Employees",
    "Employees with Disabilities"
  )

# Color matrix to color table
colors <- met.brewer("Hiroshige", 5)[c(1,4)]
colormatrix <- ifelse(median_df_wide[, -(1:2)] > median_df_wide$`All Employees`, colors[1], colors[2])

# Make formatted table
flextable(median_df_wide) |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header") |> 
  valign(valign = "center", part = "header") |>
  color(j = 3:7, color = colormatrix) 
```

Median salaries (CDN per annum) of Canadian core public service employees from 2017-2021, rounded to the nearest \$50. Median salary was estimated with [Formula 1](https://employment-equity.github.io/DI.html#estimating-median-salary). Orange values indicate a higher value than the "all employees" group, and blue values indicate a lower value per year. Data from [Treasury Board of Canada Secretariat Diversity and Inclusion Statistics.](https://www.canada.ca/en/treasury-board-secretariat/services/innovation/human-resources-statistics/diversity-inclusion-statistics.html)
:::

::: {.content-visible when-profile="french"}
## Tableau des salaires médians

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Add better group labels 
colnames(median_df_wide) <-
  c(
    "Année",
    "Tous les employés",
    "Employés Non-Minorité visible",
    "Employés Minorité visible",
    "Employés de race noire",
    "Employés autochtones",
    "Employés handicapées"
  )

# Color matrix to color table
colors <- met.brewer("Hiroshige", 5)[c(1,4)]
colormatrix <- ifelse(median_df_wide[, -(1:2)] > median_df_wide$`Tous les employés`, colors[1], colors[2]) 

# HERE
median_df_wide_fr <- median_df_wide[]


format([], round(as.numeric(), nsmall=1, big.mark=" "))
  



# Make formatted table
flextable(median_df_wide) |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header") |> 
  valign(valign = "center", part = "header") |>
  color(j = 3:7, color = colormatrix) 
```

Salaires médians (en dollars canadiens par an) des employés du noyau de la fonction publique canadienne de 2017 à 2021, arrondis aux 50 dollars les plus proches. Pour savoir comment le salaire médian a été calculé, veuillez consulter la page [Sur l'indice de disproportionnalité.](https://employment-equity.github.io/fr/DI.html) Les valeurs orange indiquent une valeur supérieure à celle du groupe "tous les salariés", et les valeurs bleues une valeur inférieure par an. Les données proviennent de [Secrétariat du Conseil du Trésor du Canada Statistiques sur la diversité et l'inclusion.](https://www.canada.ca/fr/secretariat-conseil-tresor/services/innovation/statistiques-ressources-humaines/statistiques-diversite-inclusion.html)
:::

::: {.content-visible when-profile="english"}
## Disproportionality Tables

::: panel-tabset
### 2022

```{r echo=FALSE, message=FALSE, warning=FALSE}

# In this chunk: calculate DI's and flextable 2022 data
di_df_2022 <- years_data_cat |>
  filter(Year == "2022") |>
  mutate(
    range_group = case_when(# group salary ranges into quintiles
      salary_range %in% c(
        "Under 50,000",
        "25,000 to 49,999",
        "50,000 to 54,999",
        "55,000 to 59,999"
      ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999", 
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999", 
      salary_range %in% c(
        "70,000 to 74,999", 
        "75,000 to 79,999", 
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
      ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> 
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)
  ) |>
  ungroup()


di_df_2022 <- 
  di_df_2022[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  dplyr::select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2022")

# ---- Export Table 2022
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
# REMOVE YEAR TOO
export_table_2022 <- flextable(select(di_df_2022, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2022 <- set_header_labels(export_table_2022,
  values = list(
    range_group = "Salary Range",
    per_all = "Percent",
    di_all = "DI",
    per_black = "Percent",
    di_black = "DI",
    per_vm = "Percent",
    di_vm = "DI",
    per_non_vm = "Percent",
    di_non_vm = "DI",
    per_indigenous = "Percent",
    di_indigenous = "DI",
    per_pwd = "Percent",
    di_pwd = "DI"
  )
)

export_table_2022 <-  add_header_row(
  export_table_2022,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "All Employees",
    "Black Employees",
    "All Visible Minority Employees",
    "Non-Visible Minority Employees",
    "Indigenous Employees",
    "Employees with Disabilities"
  )
)

export_table_2022 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2021

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2021 data
di_df_2021 <- years_data_cat |>  
  filter(Year == "2021") |>
  mutate(range_group = case_when(
      salary_range %in% c(
        "Under 25,000",
        "25,000 to 49,999",
        "50,000 to 54,999",
        "55,000 to 59,999"
      ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999", 
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999", 
        "75,000 to 79,999", 
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
        ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges 
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup()

di_df_2021 <- 
  di_df_2021[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  dplyr::select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2021")

# ---- Export Table 2021
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2021 <- flextable(select(di_df_2021, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2021 <- set_header_labels(export_table_2021,
  values = list(
    range_group = "Salary Range",
    per_all = "Percent",
    di_all = "DI",
    per_black = "Percent",
    di_black = "DI",
    per_vm = "Percent",
    di_vm = "DI",
    per_non_vm = "Percent",
    di_non_vm = "DI",
    per_indigenous = "Percent",
    di_indigenous = "DI",
    per_pwd = "Percent",
    di_pwd = "DI"
  )
)

export_table_2021 <-  add_header_row(
  export_table_2021,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "All Employees",
    "Black Employees",
    "All Visible Minority Employees",
    "Non-Visible Minority Employees",
    "Indigenous Employees",
    "Employees with Disabilities"
  )
)

export_table_2021 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2020

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2020 data
di_df_2020 <- years_data_cat |>  
  filter(Year == "2020") |> mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
      salary_range %in% c(
        "Under 50,000",
        "50,000 to 54,999",
        "55,000 to 59,999"
        ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999",
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
      ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges into quintiles
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups

di_df_2020 <- 
  di_df_2020[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2020")

# ---- Export Table 2020
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2020 <- flextable(select(di_df_2020, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2020 <- set_header_labels(export_table_2020,
  values = list(
    range_group = "Salary Range",
    per_all = "Percent",
    di_all = "DI",
    per_black = "Percent",
    di_black = "DI",
    per_vm = "Percent",
    di_vm = "DI",
    per_non_vm = "Percent",
    di_non_vm = "DI",
    per_indigenous = "Percent",
    di_indigenous = "DI",
    per_pwd = "Percent",
    di_pwd = "DI"
  )
)
export_table_2020 <-  add_header_row(export_table_2020,
                     colwidths = c(1, 2,2,2,2,2,2),
                     values = c(" ", "All Employees", "Black Employees", "All Visible Minority Employees", "Non-Visible Minority Employees", "Indigenous Employees", "Employees with Disabilities"))

export_table_2020 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2019

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2019 data
di_df_2019 <- years_data_cat |>  
  filter(Year == "2019") |> mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
      salary_range %in% c(
        "Under 15,000",
        "15,000 to 49,999",
        "50,000 to 54,999",
        "55,000 to 59,999"
        ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999",
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
        ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges into quintiles
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups


di_df_2019 <- 
  di_df_2019[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2019")

# ---- Export Table 2019
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2019 <- flextable(select(di_df_2019, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2019 <- set_header_labels(export_table_2019,
  values = list(
    range_group = "Salary Range",
    per_all = "Percent",
    di_all = "DI",
    per_black = "Percent",
    di_black = "DI",
    per_vm = "Percent",
    di_vm = "DI",
    per_non_vm = "Percent",
    di_non_vm = "DI",
    per_indigenous = "Percent",
    di_indigenous = "DI",
    per_pwd = "Percent",
    di_pwd = "DI"
  )
)
export_table_2019 <-  add_header_row(
  export_table_2019,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "All Employees",
    "Black Employees",
    "All Visible Minority Employees",
    "Non-Visible Minority Employees",
    "Indigenous Employees",
    "Employees with Disabilities"
  )
)

export_table_2019 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2018

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2018 data
di_df_2018 <- years_data_cat |>  
  filter(Year == "2018") |> mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
 salary_range %in% c(
   "Under 50,000",
   "50,000 to 54,999",
   "55,000 to 59,999"
   ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999",
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 and over"
        ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges to match Dr. Martin's numbers
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups


di_df_2018 <- 
  di_df_2018[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2018")


# ---- Export Table 2018
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2018 <- flextable(select(di_df_2018, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2018 <- set_header_labels(export_table_2018,
  values = list(
    range_group = "Salary Range",
    per_all = "Percent",
    di_all = "DI",
    per_black = "Percent",
    di_black = "DI",
    per_vm = "Percent",
    di_vm = "DI",
    per_non_vm = "Percent",
    di_non_vm = "DI",
    per_indigenous = "Percent",
    di_indigenous = "DI",
    per_pwd = "Percent",
    di_pwd = "DI"
  )
)
export_table_2018 <-  add_header_row(export_table_2018,
                     colwidths = c(1, 2,2,2,2,2,2),
                     values = c(" ", "All Employees", "Black Employees", "All Visible Minority Employees", "Non-Visible Minority Employees", "Indigenous Employees", "Employees with Disabilities"))


export_table_2018 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2017

```{r echo=FALSE, message=FALSE, warning=FALSE}

# In this chunk: calculate DI's and flextable 2017 data
di_df_2017 <- years_data_cat |>  
  filter(Year == "2017") |> dplyr::mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
   salary_range %in% c("Under 50,000",
                          "50,000 to 54,999"
                          ) ~ "Less than $55,000",
      salary_range %in% c("55,000 to 59,999", 
                          "60,000 to 64,999"
                          ) ~ "$55,000 to $64,999",
      salary_range %in% c("65,000 to 69,999", 
                          "70,000 to 74,999", 
                          "75,000 to 79,999"
                          ) ~ "$65,000 to $79,999",
      salary_range %in% c("80,000 to 84,999", 
                          "85,000 to 89,999",
                          "90,000 to 94,999"
                          ) ~ "$80,000 to $94,999",
      salary_range %in% c("95,000 to 99,999", 
                          "100,000 and over"
                          ) ~ "Over $95,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges into quintiles
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups

di_df_2017 <- 
  di_df_2017[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2017")

# ---- Export Table 2017
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2017 <- flextable(select(di_df_2017, -contains("chart")) |> select(-c("year")))
# Change header labels
export_table_2017 <- set_header_labels(export_table_2017,
  values = list(
    range_group = "Salary Range",
    per_all = "Percent",
    di_all = "DI",
    per_black = "Percent",
    di_black = "DI",
    per_vm = "Percent",
    di_vm = "DI",
    per_non_vm = "Percent",
    di_non_vm = "DI",
    per_indigenous = "Percent",
    di_indigenous = "DI",
    per_pwd = "Percent",
    di_pwd = "DI"
  )
)

export_table_2017 <-  add_header_row(
  export_table_2017,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "All Employees",
    "Black Employees",
    "All Visible Minority Employees",
    "Non-Visible Minority Employees",
    "Indigenous Employees",
    "Employees with Disabilities"
  )
)

export_table_2017 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```
:::
:::

::: {.content-visible when-profile="french"}
## Tableaux de disproportionnalité

::: panel-tabset
### 2022

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2022 <- flextable(select(di_df_2022, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2022 <- set_header_labels(export_table_2022,
  values = list(
    range_group = "Fourchette des salaires",
    per_all = "Pourcentage",
    di_all = "ID",
    per_black = "Pourcentage",
    di_black = "ID",
    per_vm = "Pourcentage",
    di_vm = "ID",
    per_non_vm = "Pourcentage",
    di_non_vm = "ID",
    per_indigenous = "Pourcentage",
    di_indigenous = "ID",
    per_pwd = "Pourcentage",
    di_pwd = "ID"
  )
)

export_table_2022 <-  add_header_row(
  export_table_2022,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "Tous les employés",
    "Employés de race noire",
    "Employés Minorité visible",
    "Employés Non-Minorité visible",
    "Employés autochtones",
    "Employés handicapées"
  )
)

export_table_2022 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")

```

### 2021

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2021 data
di_df_2021 <- years_data_cat |>  
  filter(Year == "2021") |>
  mutate(range_group = case_when(
      salary_range %in% c(
        "Under 25,000",
        "25,000 to 49,999",
        "50,000 to 54,999",
        "55,000 to 59,999"
      ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999", 
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999", 
        "75,000 to 79,999", 
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
        ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges 
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup()

di_df_2021 <- 
  di_df_2021[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  dplyr::select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2021")

# ---- Export Table 2021
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2021 <- flextable(select(di_df_2021, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2021 <- set_header_labels(export_table_2021,
values = list(
    range_group = "Fourchette des salaires",
    per_all = "Pourcentage",
    di_all = "ID",
    per_black = "Pourcentage",
    di_black = "ID",
    per_vm = "Pourcentage",
    di_vm = "ID",
    per_non_vm = "Pourcentage",
    di_non_vm = "ID",
    per_indigenous = "Pourcentage",
    di_indigenous = "ID",
    per_pwd = "Pourcentage",
    di_pwd = "ID"
  )
)

export_table_2021 <-  add_header_row(
  export_table_2021,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "Tous les employés",
    "Employés de race noire",
    "Employés Minorité visible",
    "Employés Non-Minorité visible",
    "Employés autochtones",
    "Employés handicapées"
  )
)

export_table_2021 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2020

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2020 data
di_df_2020 <- years_data_cat |>  
  filter(Year == "2020") |> mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
      salary_range %in% c(
        "Under 50,000",
        "50,000 to 54,999",
        "55,000 to 59,999"
        ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999",
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
      ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges into quintiles
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups

di_df_2020 <- 
  di_df_2020[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2020")

# ---- Export Table 2020
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2020 <- flextable(select(di_df_2020, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2020 <- set_header_labels(export_table_2020,
values = list(
    range_group = "Fourchette des salaires",
    per_all = "Pourcentage",
    di_all = "ID",
    per_black = "Pourcentage",
    di_black = "ID",
    per_vm = "Pourcentage",
    di_vm = "ID",
    per_non_vm = "Pourcentage",
    di_non_vm = "ID",
    per_indigenous = "Pourcentage",
    di_indigenous = "ID",
    per_pwd = "Pourcentage",
    di_pwd = "ID")
)

export_table_2020 <-  add_header_row(export_table_2020,
                     colwidths = c(1, 2,2,2,2,2,2),
  values = c(
    " ",
    "Tous les employés",
    "Employés de race noire",
    "Employés Minorité visible",
    "Employés Non-Minorité visible",
    "Employés autochtones",
    "Employés handicapées"))

export_table_2020 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2019

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2019 data
di_df_2019 <- years_data_cat |>  
  filter(Year == "2019") |> mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
      salary_range %in% c(
        "Under 15,000",
        "15,000 to 49,999",
        "50,000 to 54,999",
        "55,000 to 59,999"
        ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999",
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 to 149,999",
        "150,000 to 199,999",
        "200,000 to 249,999",
        "250,000 and over"
        ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges into quintiles
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups


di_df_2019 <- 
  di_df_2019[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2019")

# ---- Export Table 2019
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2019 <- flextable(select(di_df_2019, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2019 <- set_header_labels(export_table_2019,
values = list(
    range_group = "Fourchette des salaires",
    per_all = "Pourcentage",
    di_all = "ID",
    per_black = "Pourcentage",
    di_black = "ID",
    per_vm = "Pourcentage",
    di_vm = "ID",
    per_non_vm = "Pourcentage",
    di_non_vm = "ID",
    per_indigenous = "Pourcentage",
    di_indigenous = "ID",
    per_pwd = "Pourcentage",
    di_pwd = "ID"
  )
)
export_table_2019 <-  add_header_row(
  export_table_2019,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "Tous les employés",
    "Employés de race noire",
    "Employés Minorité visible",
    "Employés Non-Minorité visible",
    "Employés autochtones",
    "Employés handicapées"
  )
)

export_table_2019 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2018

```{r echo=FALSE, message=FALSE, warning=FALSE}
# In this chunk: calculate DI's and flextable 2018 data
di_df_2018 <- years_data_cat |>  
  filter(Year == "2018") |> mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
 salary_range %in% c(
   "Under 50,000",
   "50,000 to 54,999",
   "55,000 to 59,999"
   ) ~ "Less than $60,000",
      salary_range %in% c(
        "60,000 to 64,999",
        "65,000 to 69,999"
        ) ~ "$60,000 to $69,999",
      salary_range %in% c(
        "70,000 to 74,999",
        "75,000 to 79,999",
        "80,000 to 84,999"
        ) ~ "$70,000 to $84,999",
      salary_range %in% c(
        "85,000 to 89,999",
        "90,000 to 94,999",
        "95,000 to 99,999"
        ) ~ "$85,000 to $99,999",
      salary_range %in% c(
        "100,000 and over"
        ) ~ "Over $100,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges to match Dr. Martin's numbers
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups


di_df_2018 <- 
  di_df_2018[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2018")


# ---- Export Table 2018
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2018 <- flextable(select(di_df_2018, -contains("chart")) |> select(-c("year")))

# Change header labels
export_table_2018 <- set_header_labels(export_table_2018,
values = list(
    range_group = "Fourchette des salaires",
    per_all = "Pourcentage",
    di_all = "ID",
    per_black = "Pourcentage",
    di_black = "ID",
    per_vm = "Pourcentage",
    di_vm = "ID",
    per_non_vm = "Pourcentage",
    di_non_vm = "ID",
    per_indigenous = "Pourcentage",
    di_indigenous = "ID",
    per_pwd = "Pourcentage",
    di_pwd = "ID"
  )
)
export_table_2018 <-  add_header_row(
  export_table_2018,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "Tous les employés",
    "Employés de race noire",
    "Employés Minorité visible",
    "Employés Non-Minorité visible",
    "Employés autochtones",
    "Employés handicapées"
  )
)


export_table_2018 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```

### 2017

```{r echo=FALSE, message=FALSE, warning=FALSE}

# In this chunk: calculate DI's and flextable 2017 data
di_df_2017 <- years_data_cat |>  
  filter(Year == "2017") |> dplyr::mutate(
    salary_range = sub(
      pattern = "    ",
      replacement = " ",
      x = salary_range
    ), # remove extra spaces in salary range column
    range_group = case_when(
   salary_range %in% c("Under 50,000",
                          "50,000 to 54,999"
                          ) ~ "Less than $55,000",
      salary_range %in% c("55,000 to 59,999", 
                          "60,000 to 64,999"
                          ) ~ "$55,000 to $64,999",
      salary_range %in% c("65,000 to 69,999", 
                          "70,000 to 74,999", 
                          "75,000 to 79,999"
                          ) ~ "$65,000 to $79,999",
      salary_range %in% c("80,000 to 84,999", 
                          "85,000 to 89,999",
                          "90,000 to 94,999"
                          ) ~ "$80,000 to $94,999",
      salary_range %in% c("95,000 to 99,999", 
                          "100,000 and over"
                          ) ~ "Over $95,000K",
      TRUE ~ "All ranges"
    )
  ) |> # group salary ranges into quintiles
  dplyr::group_by(range_group) |>
  dplyr::summarize(
    count_all = sum(count_all, na.rm = TRUE),
    count_black = sum(count_black, na.rm = TRUE),
    count_vm = sum(count_vm, na.rm = TRUE),
    count_non_vm = count_all - count_vm,
    count_indigenous = sum(count_indigenous, na.rm = TRUE),
    count_pwd = sum(count_pwd, na.rm = TRUE)) |> 
  ungroup() # summarize values in predetermined salary range groups

di_df_2017 <- 
  di_df_2017[c(5, 1, 2, 3, 6, 4), ] |> # reorder rows
  dplyr::mutate(
    # create new column for percentage of representation within subgroup
    per_all = round(count_all / count_all[n()] * 100, 2),
    per_black = round(count_black / count_black[n()] * 100, 2),
    per_vm = round(count_vm / count_vm[n()] * 100, 2),
    per_non_vm = round(count_non_vm / count_non_vm[n()] * 100, 2),
    per_indigenous = round(count_indigenous / count_indigenous[n()] * 100, 2),
    per_pwd = round(count_pwd / count_pwd[n()] * 100, 2),
    
    # create new column for disproportionality index (DI)
    di_all = 1,
    di_black = round(per_black / per_all, 2),
    di_vm = round(per_vm / per_all, 2),
    di_non_vm = round(per_non_vm / per_all, 2),
    di_indigenous = round(per_indigenous / per_all, 2),
    di_pwd = round(per_pwd / per_all, 2),
    
    # create new column to highlight difference from 1 for DI
    chart_black = di_black - 1,
    chart_vm = di_vm - 1,
    chart_non_vm = di_non_vm - 1,
    chart_indigenous = di_indigenous - 1,
    chart_pwd = di_pwd - 1,
  ) |>
  select(
    range_group,
    per_all,
    di_all,
    
    per_black,
    di_black,
    chart_black,
    
    per_vm,
    di_vm,
    chart_vm,
    
    per_non_vm,
    di_non_vm,
    chart_non_vm,
    
    per_indigenous,
    di_indigenous,
    chart_indigenous,
    
    per_pwd,
    di_pwd,
    chart_pwd
  ) |> 
  mutate(year = "2017")
# ---- Export Table 2017
# Create table to show raw data 
# Remove chart columns amd convert to flextable object
export_table_2017 <- flextable(select(di_df_2017, -contains("chart")) |> select(-c("year")))
# Change header labels
export_table_2017 <- set_header_labels(export_table_2017,
values = list(
    range_group = "Fourchette des salaires",
    per_all = "Pourcentage",
    di_all = "ID",
    per_black = "Pourcentage",
    di_black = "ID",
    per_vm = "Pourcentage",
    di_vm = "ID",
    per_non_vm = "Pourcentage",
    di_non_vm = "ID",
    per_indigenous = "Pourcentage",
    di_indigenous = "ID",
    per_pwd = "Pourcentage",
    di_pwd = "ID"
  )
)

export_table_2017 <-  add_header_row(
  export_table_2017,
  colwidths = c(1, 2, 2, 2, 2, 2, 2),
  values = c(
    " ",
    "Tous les employés",
    "Employés de race noire",
    "Employés Minorité visible",
    "Employés Non-Minorité visible",
    "Employés autochtones",
    "Employés handicapées"
  )
)

export_table_2017 |>
  align(align = "center", part = "body") |>
  align(align = "center", part = "header")
```
:::
:::

```{r echo = FALSE, message=FALSE, warning=FALSE, include = FALSE}

# Function to move data to long format for DI plotting
format_DI <- function(Year) {
  assign(paste("di_df_long", Year, sep = "_"), # create an object with year
         
         {
           get(paste("di_df", Year, sep = "_")) |> # pull di_df_year df's
             pivot_longer(cols = matches(c("per", "di", "chart")),
                          names_to = "temp",
                          values_to = "value") |> # reshape data so it's easier to deal with
             mutate(
               # separate into unit of measure
               uom = case_when(
                 grepl(pattern = "^per", x = as.character(temp)) ~ "Percent",
                 grepl(pattern = "^di", x = as.character(temp)) ~ "DI",
                 grepl(pattern = "^chart", x = as.character(temp)) ~ "Chart"
               ),
               # separate into subgroups
               subgroup = case_when(
                 grepl(pattern = "all$", x = temp) ~ "All",
                 grepl(pattern = "black$", x = temp) ~ "Black",
                 grepl(pattern = "non_vm$", x = temp) ~ "Non Visible Minorities",
                 grepl(pattern = "_vm$", x = temp) ~ "Visible Minorities",
                 grepl(pattern = "indigenous$", x = temp) ~ "Indigenous",
                 grepl(pattern = "pwd$", x = temp) ~ "Persons with disabilities"
               )
             ) |>
             dplyr::select(
               "Salary range" = range_group,
               Value = value,
               "Unit of measure" = uom,
               Subgroup = subgroup
             ) |>
             filter(`Salary range` != "All ranges") |>
             filter(Subgroup != "All") |> # Remove "all" category entries
             mutate(Subgroup = factor(
               Subgroup,
               levels = c(
                 "Non Visible Minorities",
                 "Visible Minorities",
                 "Persons with disabilities",
                 "Indigenous",
                 "Black"
               ),
               labels = c(
                 "Non-Visible Minority Employees",
                 "Visible Minority Employees",
                 "Employees with Disabilities",
                 "Indigenous Employees",
                 "Black Employees"
               )
             )) |> filter(`Unit of measure` == "DI") |> # Select only DI data
             mutate(`Salary range` = factor(
               `Salary range`,
               levels = c(
                 "Less than $60,000",
                 "$60,000 to $69,999",
                 "$70,000 to $84,999",
                 "$85,000 to $99,999",
                 "Over $100,000K",
                 "Less than $55,000",
                 "$55,000 to $64,999",
                 "$65,000 to $79,999",
                 "$80,000 to $94,999",
                 "Over $95,000K"
               ),
               labels = c(
                 "<60.0K",
                 "60.0-69.9K",
                 "70.0-84.9K",
                 "85.0-99.9K",
                 "100.0K+",
                 "<55.0K",
                 "55.0-64.9K",
                 "65.0-79.9K",
                 "80.0-94.9K",
                 "95.0K+"
               )
             )) |>
             mutate(
               quintile = dplyr::recode(
                 `Salary range`,
                 "<60.0K" = "Bottom 20th Percentile",
                 "60.0-69.9K" = "20th to 40th Percentile",
                 "70.0-84.9K" = "40th to 60th Percentile",
                 "85.0-99.9K" = "60th to 80th Percentile",
                 "100.0K+" = "Top 20th Percentile",
                 "<55.0K" = "Bottom 20th Percentile",
                 "55.0-64.9K" = "20th to 40th Percentile",
                 "65.0-79.9K" = "40th to 60th Percentile",
                 "80.0-94.9K" = "60th to 80th Percentile",
                 "95.0K+" = "Top 20th Percentile"
               )
             )
           
           
         }, envir = globalenv())
  
}        



# Apply to all years
lapply(years, format_DI) |> 
  bind_rows()

# Load in Appending function
AppendMe <- function(dfNames) {
  do.call(rbind, lapply(dfNames, function(x) {
    cbind(get(x), source = x)
  }))
}

# Append salary data and keep a column with names
di_datasets <- AppendMe(paste("di_df_long", years, sep = "_"))  |>
  mutate(
    source = dplyr::recode(
      source,
      # Change id columns to years using factors
      di_df_long_2022 = "2022",
      di_df_long_2021 = "2021",
      di_df_long_2020 = "2020",
      di_df_long_2019 = "2019",
      di_df_long_2018 = "2018",
      di_df_long_2017 = "2017"
    )
  ) |>
  rename(year = source) # rename source column to year

# Create french dataset
di_datasets_fr <-
  di_datasets |>
  mutate(Subgroup = factor(
               Subgroup,
               levels = c(
                 "Non-Visible Minority Employees",
                 "Visible Minority Employees",
                 "Employees with Disabilities",
                 "Indigenous Employees",
                 "Black Employees"
               ),
               labels = c(
                 "Employés Non-Minorité visible",
                 "Employés Minorité visible",
                 "Employés handicapées",
                 "Employés autochtones",
                 "Employés de race noire"
               )
             )) |>
    mutate(quintile = factor(
               quintile,
               levels = c(
                 "Bottom 20th Percentile",
                 "20th to 40th Percentile",
                 "40th to 60th Percentile",
                 "60th to 80th Percentile",
                 "Top 20th Percentile"
               ),
               labels = c(
                 "20e inférieur",
                 "20e à 40e",
                 "40e à 60e",
                 "60e à 80e",
                 "20e supérieur"
               )
             ))

```

::: {.content-visible when-profile="english"}
## Plot 1: DI of all groups by salary range (2022)

The dotted line represents equitable representation (a DI of 1). It can been seen that Black employees are overrepresented at the lower salary ranges and underrepresented at higher ranges compared to the other groups. As can be seen by the "Non-Visible Minority Employees" bars, an equal DI across all salary ranges should be expected.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 7.5}

# Grab just 2022 data
plot1_df <-  di_datasets |> filter(year == "2022")

# Make a vector with corresponding colors
colors = met.brewer("Johnson", 5)

# Bar plot showing salary ranges by group
ggplot(plot1_df, aes(fill = Subgroup,
                            y = Value,
                            x = `Salary range`)) +
  geom_bar(position = 'dodge',
           stat ='identity') +
  theme_classic() +
  xlab("Salary range") +
  ylab("Disproportionality Index") +
  geom_hline(yintercept = 1,
             linetype = "dashed") +
  scale_fill_manual(
    values = colors) +
  coord_cartesian(ylim = c(.5, 1.5)) +
  theme(legend.position = "top",
        legend.title=element_blank(),
       text = element_text(size = 15),
       legend.text = element_text(size = 12)) +
  guides(fill = guide_legend(nrow=2, byrow=TRUE)) +
  annotate("text",
           y = 1.02,
            x= "100.0K+",
            hjust = .2,
            label = "Overrepresented") +
  annotate("text",
           y = 0.98,
            x= "100.0K+",
            hjust = .2,
            label = "Underrepresented")
```

## Plot 2: A direct comparison of DI in Black Employees vs. Non-Visible Minority Employees (2022)

This plot zooms in on our "baseline" group and compares it to the Black employee group, which appears to be the most affected by inequity. A steady decrease in representation with each increase in salary range is clearly visible for the Black employee group.

```{r Plot 2, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7.5}

# Grab data 
plot2_df <-
  di_datasets |> filter(year == "2022" &
                           Subgroup %in% c("Non-Visible Minority Employees", "Black Employees"))

# Set color palette 
colors <- met.brewer("Johnson", 5)[c(1,5)]

# Plot 2 
ggplot(plot2_df) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    aes(
      x = Subgroup,
      y = Value,
      fill = Subgroup,
      alpha = `Salary range`
    )
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15),
    legend.text = element_text(size = 12)
  ) +
  scale_fill_manual  (values = colors, ) +
  labs(title = NULL,
       x = NULL,
       y = "Disproportionality Index",
       fill = NULL) +
  guides(fill = "none") +
  geom_hline(yintercept = 1,
             linetype = "dashed") +
  annotate(
    "text",
    y = 1.02,
    x = "Black Employees",
    hjust = -.7,
    label = "Overrepresented"
  ) +
  annotate(
    "text",
    y = 0.98,
    x = "Black Employees",
    hjust = -.7,
    label = "Underrepresented"
  )
```

## Plot 3: Interactive DI Visualization

Here you can select years and groups to display. Click the drop down on the left to filter by year. Click each group's text under the subgroup heading on the right to remove and add data. Use your mouse to select a section to zoom into. You can also hover over each bars to see each individual data point.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 10}
# Wrap text on labels
plot3_df <- di_datasets |> mutate(quintile = recode(quintile,
                 "Bottom 20th Percentile" = "Bottom 20th",
                 "20th to 40th Percentile" = "20th to 40th",
                 "40th to 60th Percentile" = "40th to 60th",
                 "60th to 80th Percentile" = "60th to 80th",
                 "Top 20th Percentile" = "Top 20th"))

# Get shared data for use with crosstalk
data_shared <- SharedData$new(plot3_df)

# Make a vector with corresponding colors
colors = met.brewer("Johnson", 5)

# great plotly barplot
plot3 <- ggplotly(ggplot(data_shared, aes(fill = Subgroup,
                            y = Value,
                            x = quintile,
                            text = paste0("Year: ", year,
                          "<br>", # Create a new line in the tooltip
                          "DI: ", Value,
                          "<br>",
                          "Subgroup: ", Subgroup,
                          "<br>",
                          "Quintile: ", quintile))) +
  geom_bar(position = 'dodge',
           stat ='identity') +
  theme_classic() +
  xlab("Salary Percentile") +
  ylab("Disproportionality Index") +
  geom_hline(yintercept = 1,
             linetype = "dashed") +
  scale_fill_manual(
    values = colors) +
  coord_cartesian(ylim = c(.5, 1.5)) +
  theme(legend.position = "top",
        legend.title=element_blank(),
       text = element_text(size = 15),
       legend.text = element_text(size = 12),
       axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)),
  tooltip = "text")



bscols(widths = c(2, NA), 
       filter_select(id = "fsid", 
                     label = "Year", 
                     sharedData = data_shared, 
                     group = ~year,
                     multiple = FALSE), 
       plot3)

```

## Plot 4: A direct comparison of DI in Black Employees vs. Non-Visible Minority Employees Over All Years (2017-2022)

```{r Plot 4, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7.5}

# set color pallette 
colors <- met.brewer("Johnson", 5)[c(1,5)]

plot4_df <- di_datasets |> mutate(quintile = recode(quintile,
                 "Bottom 20th Percentile" = "Bottom 20th",
                 "20th to 40th Percentile" = "20th to 40th",
                 "40th to 60th Percentile" = "40th to 60th",
                 "60th to 80th Percentile" = "60th to 80th",
                 "Top 20th Percentile" = "Top 20th")) |> 
  filter(Subgroup %in% c("Non-Visible Minority Employees", "Black Employees"))
  

# Bar plot showing salary ranges by group
ggplot(plot4_df) +
  geom_bar(stat = "identity",
           position = "dodge",
           aes(x = year,
               y = Value,
               fill = Subgroup,
               alpha = quintile)) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15),
    legend.text = element_text(size = 12)
  ) +
  scale_fill_manual(values = colors) +
  labs(title = NULL,
       x = "Year",
       y = "Disproportionality Index") +
  coord_cartesian(ylim = c(.5, 1.5)) +
    geom_hline(yintercept = 1,
             linetype = "dashed") +
  guides(alpha=guide_legend(title="Salary Quintile"), 
         fill = guide_legend(title = "Group"))



```
:::

:::

::: {.content-visible when-profile="french"}
## Graphique 1: ID de tous les groupes par tranche de salaire (2022)

La ligne en pointillé représente la représentation équitable (ID de 1). On constate que les salariés noirs sont surreprésentés dans les fourchettes de salaires inférieures et sous-représentés dans les fourchettes supérieures par rapport aux autres groupes. Comme le montrent les barres "Salariés n'appartenant pas à une minorité visible", on devrait s'attendre à une ID égale dans toutes les fourchettes de salaires.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 7.5}

# Edit labels
# Grab just 2022 data
plot1_df <-  di_datasets_fr |> 
  filter(year == "2022")
  

# Make a vector with corresponding colors
colors = met.brewer("Johnson", 5)

# Bar plot showing salary ranges by group
ggplot(plot1_df, aes(fill = Subgroup,
                            y = Value,
                            x = `Salary range`)) +
  geom_bar(position = 'dodge',
           stat ='identity') +
  theme_classic() +
  xlab("Fourchette des salaires") +
  ylab("Indice de disproportionnalité") +
  geom_hline(yintercept = 1,
             linetype = "dashed") +
  scale_fill_manual(
    values = colors) +
  coord_cartesian(ylim = c(.5, 1.5)) +
  theme(legend.position = "top",
        legend.title=element_blank(),
       text = element_text(size = 15),
       legend.text = element_text(size = 12)) +
  guides(fill = guide_legend(nrow=2, byrow=TRUE)) +
  annotate("text",
           y = 1.02,
            x= "100.0K+",
            hjust = .2,
            label = "Surreprésentés") +
  annotate("text",
           y = 0.98,
            x= "100.0K+",
            hjust = .2,
            label = "Sous-représentés")
```

## Graphique 2: Comparaison directe de l'AI chez les employés du race noirs et chez les employés n'appartenant pas à une minorité visible (2022)

Ce graphique se concentre sur notre groupe de référence et le compare au groupe des employés noirs, qui semble être le plus touché par l'iniquité. Une diminution constante de la représentation avec chaque augmentation de l'échelle salariale est clairement visible pour le groupe des employés noirs.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7.5}

# Grab data 
plot2_df <-
  di_datasets_fr |> filter(
    year == "2022" &
      Subgroup %in% c("Employés Non-Minorité visible", "Employés de race noire")
  ) |>
  rename("Fourchette des salaires" = "Salary range")

# Set color palette 
colors <- met.brewer("Johnson", 5)[c(1,5)]

# Plot 2 
ggplot(plot2_df) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    aes(
      x = Subgroup,
      y = Value,
      fill = Subgroup,
      alpha = `Fourchette des salaires`
    )
  ) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15),
    legend.text = element_text(size = 12)
  ) +
  scale_fill_manual  (values = colors, ) +
  labs(title = NULL,
       x = NULL,
       y = "Indice de disproportionnalité",
       fill = NULL) +
  guides(fill = "none") +
  geom_hline(yintercept = 1,
             linetype = "dashed") +
  annotate(
    "text",
    y = 1.02,
    x = "Employés de race noire",
    hjust = -.7,
    label = "Surreprésentés"
  ) +
  annotate(
    "text",
    y = 0.98,
    x = "Employés de race noire",
    hjust = -.7,
    label = "Sous-représentés"
  )
```

## Graphique 3: Visualisation interactive de l'ID

Vous pouvez ici sélectionner les années et les groupes à afficher. Cliquez sur le menu déroulant à gauche pour filtrer par année. Cliquez sur le texte de chaque groupe sous l'intitulé du sous-groupe à droite pour supprimer et ajouter des données. Utilisez votre souris pour sélectionner une section à agrandir. Vous pouvez également passer la souris sur chaque barre pour voir chaque point de données individuel.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 10}
# Wrap text on labels
plot3_df <- di_datasets_fr |>
  rename("Sous-group" = "Subgroup")

# Get shared data for use with crosstalk
data_shared <- SharedData$new(plot3_df)

# Make a vector with corresponding colors
colors = met.brewer("Johnson", 5)

# great plotly barplot
plot3 <- ggplotly(ggplot(data_shared, aes(fill = `Sous-group`,
                            y = Value,
                            x = quintile,
                            text = paste0("Année: ", year,
                          "<br>", # Create a new line in the tooltip
                          "DI: ", Value,
                          "<br>",
                          "Sous-groupe: ", `Sous-group`,
                          "<br>",
                          "Quintile: ", quintile))) +
  geom_bar(position = 'dodge',
           stat ='identity') +
  theme_classic() +
  xlab("Percentile de salaire") +
  ylab("Indice de disproportionnalité") +
  geom_hline(yintercept = 1,
             linetype = "dashed") +
  scale_fill_manual(
    values = colors) +
  coord_cartesian(ylim = c(.5, 1.5)) +
  theme(legend.position = "top",
        legend.title=element_blank(),
       text = element_text(size = 15),
       legend.text = element_text(size = 12),
       axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)),
  tooltip = "text")



bscols(widths = c(2, NA), 
       filter_select(id = "fsid", 
                     label = "Année", 
                     sharedData = data_shared, 
                     group = ~year,
                     multiple = FALSE), 
       plot3)

```

## Graphique 4: Comparaison directe des DI des employés noirs par rapport aux employés n'appartenant pas à une minorité visible sur toutes les années. (2017-2022)

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 7.5}

# set color pallette 
colors <- met.brewer("Johnson", 5)[c(1,5)]

plot4_df <- di_datasets_fr |> 
  filter(Subgroup %in% c("Employés Non-Minorité visible", "Employés de race noire"))
  

# Bar plot showing salary ranges by group
ggplot(plot4_df) +
  geom_bar(stat = "identity",
           position = "dodge",
           aes(x = year,
               y = Value,
               fill = Subgroup,
               alpha = quintile)) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    text = element_text(size = 15),
    legend.text = element_text(size = 12)
  ) +
  scale_fill_manual(values = colors) +
  labs(title = NULL,
       x = "Année",
       y = "Indice de disproportionnalité") +
  coord_cartesian(ylim = c(.5, 1.5)) +
    geom_hline(yintercept = 1,
             linetype = "dashed") +
  guides(alpha=guide_legend(title="Quintile de salaire"), 
         fill = guide_legend(title = "Sous-group"))


```
:::

:::
